\section{Related Work}
\label{sec:related}

\paragraph{Vision--language--action models.}
OpenVLA~\citep{kim2024openvla} extends a discrete tokenisation paradigm to
embodied control by training a 7B Llama-2 to autoregressively emit action
tokens. Latency at inference forced subsequent designs toward continuous
heads: $\pi_0$~\citep{black2024pi0} introduces a flow-matching action expert
that conditions on a frozen PaliGemma backbone, and
$\pi_{0.5}$~\citep{black2025pi05} adds knowledge insulation---blocking
gradients from the action expert from corrupting the backbone---together with
adaRMS-injected flow time, three multi-view image inputs and a discretised
state encoded into the language prompt. Our work treats the
$\pi_{0.5}$ knowledge-insulated interface as the explicit object of analysis.

\paragraph{Sparse-autoencoder dictionaries.}
\citet{cunningham2023sparse} and \citet{bricken2023monosemanticity} demonstrate
that an overcomplete dictionary trained with an L1 penalty extracts
near-monosemantic features from transformer LLMs.
\citet{gao2024scaling} replace L1 with a hard top-$k$ activation, fixing
sparsity exactly and enabling stable training at scale; we follow their recipe
including the auxiliary dead-feature reanimation loss. Closest to our setting,
\citet{swann2026sae} train SAEs on the residual stream of OpenVLA-style
discrete VLAs and show that learned features are interpretable and
\emph{steerable}---perturbations along single dictionary directions
predictably shift the policy's action distribution. Our experiments target the
continuous flow-matching expert of $\pi_{0.5}$ rather than a discrete decoder
and study how the saturating geometry of normalised actions changes the
steering response.

\paragraph{Flow matching for control.}
Flow matching~\citep{lipman2023flow} learns a continuous velocity field whose
integration transports noise to data. In robotics it offers a single-pass
alternative to denoising diffusion policies. Concurrent
work~\citep{black2025pi05} reports that the flow expert benefits from a
\emph{knowledge insulation} interface that prevents continuous gradients from
degrading the backbone; we provide a first internal characterisation of this
interface using sparse-coding tools.
