\section{Reproducibility details}
\label{app:repro}

\paragraph{Model.}
\texttt{lerobot/pi05\_libero\_base} (HuggingFace), float32, $3.6$B parameters.
We instantiate it through \texttt{lerobot.common.policies.pi05.modeling\_pi05.PI05Policy}
with the LIBERO-10 \texttt{stats.json} as dataset normalisation statistics
and the public PaliGemma SentencePiece tokenizer pulled from
\texttt{gs://big\_vision/paligemma\_tokenizer.model}.

\paragraph{Hook implementation.}
We register a forward hook on \texttt{policy.model.action\_out\_proj} that
captures \texttt{inp[0]} (shape $[B,H,d_e]{=}[1,10,1024]$) at every flow-matching
denoise step, and additionally wrap \texttt{paligemma\_with\_expert.forward} to
log final prefix and suffix outputs. The wrapping is required because
\texttt{PaliGemmaWithExpertModel} runs its layers inside a Python loop with
local rebinding of \texttt{inputs\_embeds}, so a standard
\texttt{register\_forward\_hook} on individual transformer blocks does not fire.

\paragraph{Steering implementation.}
\texttt{torch.nn.Module.register\_forward\_pre\_hook} on
\texttt{action\_out\_proj} replaces the input with $x + \alpha\,W_d[j]$
on every denoise step. Same observation and noise tensor are reused across
$(\alpha,j)$ pairs for paired comparisons.

\paragraph{Compute.}
A single A100-80GB. End-to-end pipeline (extract $480$ frames $\rightarrow$
train SAE $50$ epochs $\rightarrow$ feature analysis $\rightarrow$ $3$
features $\times\,5$ $\alpha$ values steering with $4$ rollouts each)
runs in $\sim$$13$ minutes on the trained model.

\paragraph{Code.}
Open-sourced at the project repository; the four scripts
\texttt{extract\_real.py}, \texttt{sae.py}, \texttt{analyze\_features.py},
\texttt{steer\_real.py} reproduce all numbers in the main body and
Fig.~\ref{fig:fig2}.
